---
title: 'Session 10: Data Science Capstone Project'
author: "Tolga Tezcan"
date: "`r Sys.Date()`"
output: 
    html_document:
      number_sections: true
      highlight: haddock
      theme: spacelab
      toc: yes
      toc_depth: 2
      toc_float:
        collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style>
div.navy1 { background-color:#686868; border-radius: 5px; padding: 20px; border-style: groove; color: #ffffff;}

</style>



```{r, load_libraries, include = FALSE}

if(!is.element("tidyverse", installed.packages()[,1]))
{  install.packages("tidyverse")}

if(!is.element("Hmisc", installed.packages()[,1]))
{  install.packages("Hmisc")} #package for data summary using `describe`

if(!is.element("ggplot2", installed.packages()[,1]))
{  install.packages("ggplot2")} #package for plots
if(!is.element("ggthemes", installed.packages()[,1]))
{  install.packages("ggthemes")} #package to make fancier ggplots

if(!is.element("janitor", installed.packages()[,1]))
{ install.packages("janitor")} #package to visualize results of machine learning tools
if(!is.element("rpart.plot", installed.packages()[,1]))
{  install.packages("rpart.plot")} #package to visualize trees

library(rpart.plot)
library(caret)
library(tidyverse) # the usual stuff: dplyr, readr, and other goodies
library(lubridate)
library(janitor) # clean_names()
library(Hmisc)
```

# Introduction and learning objectives

<div class = "navy1">
The purpose of this exercise is to build an estimation engine to guide investment decisions in London house market. You will first build machine learning algorithms (and tune them) to estimate the house prices given variety of information about each property. Then, using your algorithm, you will choose 200 houses to invest in out of about 2000 houses on the market at the moment.


<b>Learning objectives</b>
 
<ol type="i">
  <li>Using different data mining algorithms for prediction.</li>
  <li>Dealing with large data sets</li>
  <li>Tuning data mining algorithms</li>
  <li>Interpreting data mining algorithms and deducing importance of variables</li>
  <li>Using results of data mining algorithms to make business decisions</li>
</ol>  
</div>

# Load data

There are two sets of data, i) training data that has the actual prices ii) out of sample data that has the asking prices. Load both data sets. 

Make sure you understand what information each column contains. Note that not all information provided might be useful in predicting house prices, but do not make any assumptions before you decide what information you use in your prediction algorithms.

```{r read-investigate}
#read in the data

london_house_prices_2019_training<-read.csv("training_data_assignment_with_prices.csv")
london_house_prices_2019_out_of_sample<-read.csv("test_data_assignment.csv")



#fix data types in both data sets

#fix dates
london_house_prices_2019_training <- london_house_prices_2019_training %>% mutate(date=as.Date(date))
london_house_prices_2019_out_of_sample<-london_house_prices_2019_out_of_sample %>% mutate(date=as.Date(date))
#change characters to factors
london_house_prices_2019_training <- london_house_prices_2019_training %>% mutate_if(is.character,as.factor)
london_house_prices_2019_out_of_sample<-london_house_prices_2019_out_of_sample %>% mutate_if(is.character,as.factor)

#take a quick look at what's in the data
str(london_house_prices_2019_training)
str(london_house_prices_2019_out_of_sample)



```


```{r split the price data to training and testing}
#let's do the initial split
library(rsample)
train_test_split <- initial_split(london_house_prices_2019_training, prop = 0.75) #training set contains 75% of the data
# Create the training dataset
train_data <- training(train_test_split)
test_data <- testing(train_test_split)



```


# Visualize data 

Visualize and examine the data. What plots could be useful here? What do you learn from these visualizations?

```{r visualize}




```

Estimate a correlation table between prices and other continuous variables. What do you glean from the correlation table?

```{r, correlation table, warning=FALSE, message=FALSE}

# produce a correlation table using GGally::ggcor()
# this takes a while to plot

library("GGally")
london_house_prices_2019_training %>% 
  select(-ID) %>% #keep Y variable last
  ggcorr(method = c("pairwise", "pearson"), layout.exp = 2,label_round=2, label = TRUE,label_size = 2,hjust = 1,nbreaks = 5,size = 2,angle = -20)

```


# Fit a linear regression model

To help you get started I build a linear regression model below. I chose a subset of the features with no particular goal. You can (and should) add more variables and/or choose variable selection methods if you want.

```{r LR model}


#Define control variables
control <- trainControl (
    method="cv",
    number=5,
    verboseIter=TRUE) #by setting this to true the model will report its progress after each estimation

#we are going to train the model and report the results using k-fold cross validation
model1_lm<-train(
    price ~ distance_to_station +water_company+property_type+whether_old_or_new+freehold_or_leasehold+latitude+ longitude,
    train_data,
   method = "lm",
    trControl = control
   )

# summary of the results
summary(model1_lm)
```


```{r}
# we can check variable importance as well
importance <- varImp(model1_lm, scale=TRUE)
plot(importance)


```

## Predict the values in testing and out of sample data

Below I use the predict function to test the performance of the model in testing data and summarize the performance of the linear regression model. How can you measure the quality of your predictions?

```{r}
# We can predict the testing values

predictions <- predict(model1_lm,test_data)

lr_results<-data.frame(  RMSE = RMSE(predictions, test_data$price), 
                            Rsquare = R2(predictions, test_data$price))

                            
lr_results                         

#We can predict prices for out of sample data the same way
predictions_oos <- predict(model1_lm,london_house_prices_2019_out_of_sample)
```

# Fit a tree model

Next I fit a tree model using the same subset of features. Again you can (and should) add more variables and tune the parameter of your tree to find a better fit. 

Compare the performance of the linear regression model with the tree model; which one performs better? Why do you think that is the case?

```{r tree model}

model2_tree <- train(
  price ~ distance_to_station +water_company+property_type+whether_old_or_new+latitude+ longitude,
  train_data,
  method = "rpart",
  trControl = control,
  tuneLength=10
    )

#You can view how the tree performs
model2_tree$results

#You can view the final tree
rpart.plot(model2_tree$finalModel)

#you can also visualize the variable importance
importance <- varImp(model2_tree, scale=TRUE)
plot(importance)

```


# Other algorithms

Use at least two other algorithms to predict prices. Don't forget to tune the parameters of these algorithms. And then compare the performances of your algorithms to linear regression and trees.

# Stacking

Use stacking to ensemble your algorithms.

```{r,warning=FALSE,  message=FALSE }


```


# Pick investments

In this section you should use the best algorithm you identified to choose 200 properties from the out of sample data.

```{r,warning=FALSE,  message=FALSE }


numchoose=200

oos<-london_house_prices_2019_out_of_sample

#predict the value of houses
oos$predict <- predict(model2_tree,oos)
#Choose the ones you want to invest here
#Make sure you choose exactly 200 of them

#output your choices. Change the name of the file to your "lastname_firstname.csv"
write.csv(oos,"my_submission.csv")

```
